Equalizer 2.0\\Convergence of a Parallel Rendering Framework

### 2017-04-03 paperIEEE.tex
More benchmarks
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@34 db7c848b-7755-4df2-b763-8796797afa52

Caching retains instance data of objects in a cache, and reuses this data to
accelarate mapping of objects. The instance cache is either filled by "snooping"
on multicast transmissions or by an explicit preloading when the master objects
are registered. The preloading ("send on register") sends instance data of
recently registered master objects while the correspinding node is idle to all
connected nodes. These nodes simply enter the received data to their
cache. Preloading uses multicast when available.

@@

RTT Scale uses a master-slave execution mode, were a single running Deltagen
instance can go into "Scale mode" at any time by launching an Equalizer
configuration. Consequently, the whole internal representation needed for
rendering is based on a Collage-based data distribution. The rendering clients
are separate, smaller applications which will map their scenes during
startup. At runtime, any change performed in the main application is committed
as a delta at the beginning of the next frame, following a design pattern
similar to the Collage \textsf{Serializable}
(\sref{sec:Serializable}). Multicast (\sref{sec:RSP}) is used to keep data
distribution times during session launch reasonable for larger cluster sizes
(tens to hundreds of nodes).

@@

\begin{figure}[ht]\center
  \includegraphics[height=.44\columnwidth]{images/spikes512}\hfil
  \includegraphics[height=.44\columnwidth]{images/beechnut}
  \caption{\label{fVolumes}Volumes used for object distribution benchmarks:
    spike frequencies of a three million neuron electrical simulation at
    437x240x512 unsigned byte (51MB, left) and an MicroCT scan of a beechnut at
    1024x1024x1546 unsigned short (3GB, right) resolution}
\end{figure}

@@

RLE compression has a very low overhead, but does not compress the data will
since it only removes "blank space" in the data. The Snappy compression, used as
default in Collage, achieves the same compression ratio as the LZ variants at a
much higher speed. The ZStandard compressor has roughly the same speed as the LZ
variants at the lowest compression level, but provides significantly better
compression. At higher compression levels it can improve the compression ratio
slightly, but at a high cost for the compression speed.

@@

In this section we present results on how data distribution and synchronization
performs in real-world applications. We extracted the existing data distribution
code in a PLY and a volume renderer into a benchmark application to measure the
time to initially map all the objects on the render client nodes, and to perform
a commit-sync of the full data set after mapping has been established. All
figures observe a noticable measurement jitter due to other services running on
the cluster. The details of the benchmark algorithm can be found in the implementation~\cite{eqObjectBench}.

David 2mm

>>>>

Caching retains instance data of objects in a cache, and reuses this data to
accelarate mapping of objects. The instance cache is either filled by
``snooping'' on multicast transmissions or by an explicit preloading when the
master objects are registered. The preloading (``send on register'') sends
instance data of recently registered master objects while the correspinding node
is idle to all connected nodes. These nodes simply enter the received data to
their cache. Preloading uses multicast when available.

@@

RTT Scale uses a master-slave execution mode, were a single running Deltagen
instance can go into ``Scale mode'' at any time by launching an Equalizer
configuration. Consequently, the whole internal representation needed for
rendering is based on a Collage-based data distribution. The rendering clients
are separate, smaller applications which will map their scenes during
startup. At runtime, any change performed in the main application is committed
as a delta at the beginning of the next frame, following a design pattern
similar to the Collage \textsf{Serializable}
(\sref{sec:Serializable}). Multicast (\sref{sec:RSP}) is used to keep data
distribution times during session launch reasonable for larger cluster sizes
(tens to hundreds of nodes).

@@

\begin{figure}[ht]\center
  \includegraphics[height=.44\columnwidth]{images/spikes512}\hfil
  \includegraphics[height=.44\columnwidth]{images/beechnut}
  \caption{\label{fVolumes}Volumes used for object distribution benchmarks:
    spike frequencies of a three million neuron electrical simulation at
    512x437x240 unsigned byte (51MB, left) and an MicroCT scan of a beechnut at
    1024x1024x1546 unsigned short (3GB, right) resolution}
\end{figure}

@@

RLE compression has a very low overhead, but does not compress the data will
since it only removes ``blank space'' in the data. The Snappy compression, used
as default in Collage, achieves the same compression ratio as the LZ variants at
a much higher speed. The ZStandard compressor has roughly the same speed as the
LZ variants at the lowest compression level, but provides significantly better
compression. At higher compression levels it can improve the compression ratio
slightly, but at a high cost for the compression speed.

@@

In this section we present how data distribution and synchronization performs in
real-world applications. We extracted the existing data distribution code in a
PLY and a volume renderer into a benchmark application to measure the time to
initially map all the objects on the render client nodes, and to perform a
commit-sync of the full data set after mapping has been established. All figures
observe a noticable measurement jitter due to other services running on the
cluster. The details of the benchmark algorithm can be found in the
implementation~\cite{eqObjectBench}. We used the same data sets as in the
previous section. After $8$ processes, a nodes start to run two processes per
node, which share the CPU and network interface.

Object mapping is measured using the following settings: \textbf{none}
distributes the raw, uncompressed, and unbuffered data, \textbf{compression}
uses the Snappy compresser to distribute unbuffered data, \textbf{buffered}
reuses uncompressed, serialized data for mappings of other nodes, and
\textbf{compression buffered} compresses this buffer. Unbuffered operations need
to reserialize, and potentially recompress, the master object data for each
slave node. Each slave instance needs to deserialize and decompress the data.

During data synchronization, the master instance commits the object data to all
mapped slave instances simultaneously. This is a ``push'' operation, whereas the
mapping is a slave-triggered ``pull'' operation with additional synchronization
points. Slave nodes queue this data and consume it during synchronization. We
test the time to commit and sync the data using different compression engines.

The David statue at 2mm resolution is organized in a kD-tree for rendering. Each
node is a separate distributable object, which has two child node objects. A
total of $1023$ objects is distributed and synchronized.

\fig{fSpikesDist} has a significant compression ratio, which reflects in the
results. Futhermore, only a single distributed object is used to transmit the
volume information in one chunk. Compression for this data is benefitial for
transmitting data over a 10Gb/s link (880 MB/s vs 660MB/s raw data
transfer). Buffering has little benefit since the serialization of volume data
is trivial. Buffered compression makes a huge difference, since the compression
cost can be amortized over $n$ nodes, reaching raw data transmission rates of
3.7 GB/s.
